---
title: "Programming assignment 3"
format: gfm
date: last-modified
date-format: iso
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| echo: false
library(tidyverse)
library(dplyr)
library(devtools)
library(untidydata)
library(readr)
library(ggplot2)
```

```{r}
#| label: data
dat <- read_csv("./data/vowel_data.csv")

```

```{r}
#| label: means and sd
means <- dat %>%
  group_by(vowel, language) %>%
  summarise(f1_cent_mean = mean(f1_cent), f1_sd = sd(f1_cent), f2_cent_mean = mean(f2_cent), f2_sd = sd(f2_cent), tl_mean = mean(tl), tl_sd = sd(tl))

```

```{r}
#| label: plots

# Trajectory length as a function of vowel anf language

dat %>%
  ggplot(aes(x = vowel, y = tl, fill = vowel)) +
  geom_boxplot() +
  facet_wrap(~ language, scales = "free_y") +
  geom_boxplot() +
  labs(x = "Vowel", y = "Trajectory length", fill = "Vowel")

# F1 as a function of vowel and language

dat %>%
  ggplot(aes(x = vowel, y = f1_cent, fill = vowel)) +
  geom_boxplot() +
  facet_wrap(~ language, scales = "free_y") +
  geom_boxplot() +
  labs(x = "Vowel", y = "F1 Cent", fill = "Vowel")

# F2 as a function of vowel and language

dat %>%
  ggplot(aes(x = vowel, y = f2_cent, fill = vowel)) +
  geom_boxplot() +
  facet_wrap(~ language, scales = "free_y") +
  geom_boxplot() +
  labs(x = "Vowel", y = "F2 Cent", fill = "Vowel")

```


```{r}

dat_long <- dat %>%
  pivot_longer(cols = f1_cent:tl, names_to = "measure", values_to = "val") %>%
  #group_by(language, measure) %>%
  #summarise(avg = mean(val)) %>%
  ggplot(aes(x = measure, y = val, fill = as.factor(vowel))) +
  geom_point() +
  facet_wrap(~ measure, scales = "free_y") +
  geom_point() +
  labs(x = "Measure", y = "Average Value", fill = "Vowel")


```


```{r}

#| label: Spectral centroids of English/Spanish cardinal vowels
#| fig-retina: 2
#| fig-height: 5
#| out-width: "100%"
#| eval: true

vowel_means <- dat |> 
  group_by(vowel, language) |> 
  summarize(f1_cent = mean(f1_cent), f2_cent = mean(f2_cent)) |> 
  ungroup() |> 
  mutate(order = case_when(vowel == "i" ~ 1, vowel == "a" ~ 2, TRUE ~ 3), 
         vowel = forcats::fct_reorder2(vowel, vowel, order)) |> 
  arrange(order)

dat |> 
  mutate(vowel = forcats::fct_relevel(vowel, "u", "a", "i")) |> 
  ggplot() +
  aes(x = f2_cent, y = f1_cent, color = language, label = vowel) + 
  geom_text(size = 3.5, alpha = 0.6, show.legend = F) + 
  geom_path(
    data = vowel_means, 
    aes(group = language, lty = language), 
    color = "grey"
  ) + 
  geom_text(data = vowel_means, show.legend = F, size = 7) + 
  scale_y_reverse() + 
  scale_x_reverse() + 
  scale_color_brewer(palette = "Set1") + 
  labs(
    title = "Vowel space comparison", 
    subtitle = "Spectral centroids of English/Spanish cardinal vowels", 
    y = "F1 (hz)", 
    x = "F2 (hz)"
  ) + 
  theme_minimal(base_size = 16)
```


# Answers to script questions

**A. Examine the portion of the script you see below. In your own words what does this section do and why does it work? Demonstrate that you understand the code…**

vonset = Get starting point: 2, 2
voffset = Get end point: 2, 2
durationV = voffset - vonset
per20 = vonset + (durationV * 0.20)
per35 = vonset + (durationV * 0.35)
per50 = vonset + (durationV * 0.50)
per65 = vonset + (durationV * 0.65)
per80 = vonset + (durationV * 0.80)

What this code is doing is calculating the percentages of a duration between two points. First, 'vonset' and 'voffset' establish the start and the endpoint of a given space. In this case, and based on what appears in the dataset, it marks the start and the endpoint of the vowel. The, it calculates the duration of that vowel by calculating the differences between the start and the ending point. Finally, it calculates certain percentages of the duration. 

**B. In a few short sentences describe the general outline of the script, what the purpose is, and how it achieves this purpose (hint: focus on the section dividers and the comments).**

The first thing the script does it to set up the output file. Essentially, it is telling Praat to create a .csv file with a specific number of columns (and the names of these columns) and to save it in a specific location (path). Then, the loop starts, which is what is going to get the data from the sound file (and the textgrid) that we have in Praat and put it in the .csv file that it asked Praat to create in the previous step. In this step, it is important to add the path to the stimuli files, that is, we have to tell the script t where to look for the sound files to get the stimuli from. Finally, the third step is to run the loop.

**C. In a few short sentences describe how the segmenting procedure you used this week differs from that used in pa_2. What are the advantages and disadvantages?**

The main difference between the code used for pa_2 and pa_3 is that the first one is divided into two separate scripts and the second one is just one long script. Additionally, their structure and what they do are a little different. The first portion of the script used in pa_2 is meant to segment the master sound file into smaller sound files. The script used for pa_3, however, does not do that. I believe this could be a disadvantage for the script used in pa_3, given that it is useful to segment the master sound file into smaller files for other purposes. The first portion of the script used in pa_3, on the other hand, is used to let the script know where and how to save the data, which is something that the script for pa_2 also has but at the very bottom. I believe it is more useful to have it at the begging, so that you have a clear idea of what where you are going to save the data and how. After this first portion, both scripts set the path to the folder where the sound files are. The script for pa_2 is going to look for several segmented files, whereas the script for pa_3 is going to look for just one file, the master file. After that, the script for pa_2 does what the script for pa_3 did at the very begging, letting the script know where and how to save the data. After that, both scripts prepare and start/run the loop












